---
title: "{{< var title >}}: {{< var version >}}"
subtitle: "EDA - proficiency/use"
author: "{{< var author.name >}}"
date: "`r Sys.Date()`"
format: 
  html: default
  pdf: default
number-sections: false
execute: 
  echo: true
  warning: false
  message: false
  fig.retina: 2
---

### Setup

Load libraries and helper functions: 

```{r}
#| label: load-packages
source(here::here("scripts", "00_libraries.R"))
source(here::here("scripts", "01_helpers.R"))
```

Load proficiency and use data: 

```{r}
#| label: load-prof-use-data
prof_use_temp <- read_xlsx(
  here("data", "raw",  "HS_L2_used_blc_V2.xlsx")
  ) %>% 
  rename(id = Participant, prof = dele, use = `span_week_%`)
```

Next, we tidy the data and convert to z-scores.

```{r}
#| label: tidy-clean-up
# Create heritage and L2 groups 
# Standardize proficiency
prof_use <- prof_use_temp %>% 
  mutate(group = case_when(
    Type == "advanced_heritage" ~ "hl", 
    Type == "intermediate_heritage" ~ "hl", 
    TRUE ~ "l2"), 
    prof_p = prof / 56, 
    use_p = use / 100, 
    prof_perc = prof_p * 100, 
    use_perc = use_p * 100, 
    prof_z = (prof_perc - mean(prof_perc)) / sd(prof_perc), 
    use_z = (use_perc - mean(use_perc)) / sd(use_perc)
  )
  
```

Basic plot to check distributions. 

```{r}
#| label: eda-prof-use-plots

prof_use %>% 
  ggplot(., aes(x = use_p, y = prof_p, color = group)) + 
    geom_point() + 
    scale_color_viridis_d(option = "A", begin = 0.3, end = 0.85)
```

The data look pretty evenly distributed, though perhaps there is more variability in the L2 group. 
Now we will look at some descriptives in order to get a sense of whether or not there are any individuals that might make the groups uneven.

```{r}
#| label: prof-use-descriptives
#| results: 'hold'
# Check mean (SD) for each TYPE
prof_use %>% 
  group_by(Type) %>% 
  summarize(across(c("prof_p", "use_p"), 
    list(mean = mean, sd = sd)))

# Might be able to balance groups by removing intermediate 
# heritage with high use
prof_use %>% 
  filter(Type == "intermediate_heritage") %>% 
  arrange(desc(use_p)) %>% 
  select(id, prof_p, use_p)

# There are three in the 70+ % range
# We'll start there
prof_use %>% 
  filter(!(id %in% c("ihs24", "ihs13", "ihs18"))) %>% 
  group_by(Type) %>% 
  summarize(across(c("prof_p", "use_p"), 
    list(mean = mean, sd = sd)))

# They are closer now. We can also take out a few of the lower L2 use
prof_use %>% 
  filter(Type == "intermediate_l2") %>% 
  arrange(use_p) %>% 
  select(id, prof_p, use_p)

# There are three in the 15% range
# We'll start there
prof_use %>% 
  filter(!(id %in% c("ihs24", "ihs13", "ihs18")), 
         !(id %in% c("ies09", "ies26", "ies29"))) %>% 
  group_by(Type) %>% 
  summarize(across(c("prof_p", "use_p"), 
    list(mean = mean, sd = sd)))

prof_use %>% 
  filter(Type == "advanced_heritage") %>% 
  arrange(desc(use_p)) %>% 
  select(id, prof_p, use_p)


prof_use %>%   filter(!(id %in% c("ihs24", "ihs13", "ihs18")), 
         !(id %in% c("ahs12", "ahs03")), 
         !(id %in% c("ies09", "ies26", "ies29"))) %>% 
  group_by(group) %>% 
  summarize(across(c("prof_p", "use_p"), 
    list(mean = mean, sd = sd)))

# IDs to drop
remove <- c(
  "ihs24", "ihs13", "ihs18", 
  "ahs12", "ahs03", 
  "ies09", "ies26", "ies29"
  )

prof_use_final <- prof_use %>% 
  filter(!(id %in% remove)) %>% 
  mutate(prof_z = (prof_perc - mean(prof_perc)) / sd(prof_perc), 
    use_z = (use_perc - mean(use_perc)) / sd(use_perc))

```

We now have a few candidates. 
We will do a few formal tests to look at equivalence in proficiency and use with and without them. 

```{r}
#| label: eda-prof-use-models

#
# Fit models
#

# model formula
model_formula <- bf(
  prop ~ group, 
  phi  ~ group 
  )

# Fit proficiency model with all data
prof_with <- brm(
  formula = model_formula, 
  family = Beta(), 
  data = mutate(prof_use, prop = prof / 56), 
  file = here("models", "eda_prof_use", "prof_with")
  )

# Fit use model with all data
use_with <- brm(
  formula = model_formula, 
  family = Beta(), 
  data = mutate(prof_use, prop = use / 100), 
  file = here("models", "eda_prof_use", "use_with")
  )

# Fit proficiency model remove a few participants
prof_wo <- brm(
  formula = model_formula, 
  family = Beta(), 
  data = mutate(prof_use_final, prop = prof / 56), 
  file = here("models", "eda_prof_use", "prof_wo")
  )

# Fit use model remove a few participants
use_wo <- brm(
  formula = model_formula, 
  family = Beta(), 
  data = mutate(prof_use_final, prop = use / 100), 
  file = here("models", "eda_prof_use", "use_wo")
  )
```

Now we can check the marginal mean estimates to see how the groups compare. 

```{r}
#| label: eda-prof-use-marginal-mean-estimates

# Get marginal mean estimates for prof with ids removed
prof_contrast <- prof_wo %>% 
  emmeans(spec = ~ group, 
    at = list(group = c("l2", "hl")), 
    epred = TRUE) %>% 
  contrast(method = "pairwise") %>% 
  gather_emmeans_draws() %>% 
  mutate(metric = "prof")

# Get marginal mean estimates for use with ids removed
use_contrast <- use_wo %>% 
  emmeans(spec = ~ group, 
    at = list(group = c("l2", "hl")), 
    epred = TRUE) %>% 
  contrast(method = "pairwise") %>% 
  gather_emmeans_draws() %>% 
  mutate(metric = "use")

use_prof_desc <- bind_rows(
  bind_cols(x = 0.15, y = 1.2, 
    metric = "Proficiency", 
    median_hdi(prof_contrast, .value), 
    ROPE = rope(prof_contrast$.value, ci_method = "HDI")$ROPE_Percentage,
    PD = p_direction(prof_contrast$.value)[[1]]
    ), 
  bind_cols(x = -0.3, y = 2.2, 
    metric = "Use", 
    median_hdi(use_contrast, .value), 
    ROPE = rope(use_contrast$.value, ci_method = "HDI")$ROPE_Percentage,
    PD = p_direction(use_contrast$.value)[[1]]
    )
  ) %>% 
  mutate(across(c(".value", ".lower", ".upper", "PD"), 
    ~ round(.x, digits = 2))) %>% 
  transmute(x, y, metric, contrast, 
    out = glue("β = {.value} [{.lower}, {.upper}]\nROPE = {ROPE}, PD = {PD}"))

bind_rows(prof_contrast, use_contrast) %>% 
  mutate(ylab = if_else(metric == "use", "Use", "Proficiency")) %>% 
  ggplot(., aes(x = .value, y = ylab, fill = stat(abs(x) > 0.1))) + 
    geom_tile(data = tibble(
      ylab = "Proficiency", value = 0, lower = -0.1, upper = 0.1
      ),
      aes(x = value, width = lower - upper),
      alpha = 0.1, height = Inf, fill = "#31688EFF") +
    geom_vline(xintercept = 0, lty = 3, size = 0.5) + 
    stat_dotsinterval(quantiles = 100, point_interval = "median_hdi", 
      show.legend = F, pch = 21, point_fill = "white") +
    scale_fill_manual(values = c("grey", "#cc0033")) + 
    coord_cartesian(xlim = c(-0.3, 0.3)) + 
    geom_text(data = use_prof_desc, aes(x = x, y = y, label = out), 
      hjust = 0, size = 3) +
    labs(y = NULL, x = "Marginal mean difference\n(L2 - HS)") + 
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5))
```

```{r}
#| label: eda-prof-use-final-descriptives
n_groups <- prof_use_final %>% 
  mutate(group = str_replace_all(group, "hl", "HS")) %>% 
  group_by(group) %>% 
  summarize(n = n_distinct(id))

tab_prof_use_desc <- prof_use_final %>% 
  mutate(group = str_replace_all(group, "hl", "HS")) %>% 
  group_by(group) %>% 
  summarize(across(c("prof_p", "use_p"), 
    list(mean = mean, sd = sd))) %>% 
  mutate_if(is.numeric, specify_decimal, k = 2) %>% 
  left_join(., n_groups, by = "group") %>% 
  mutate(
    Proficiency = glue("{prof_p_mean} ({prof_p_sd})"), 
    Proficiency_prose = glue("{prof_p_mean} (SD = {prof_p_sd})"), 
    Use = glue("{use_p_mean} ({use_p_sd})"), 
    Use_prose = glue("{use_p_mean} (SD = {use_p_sd})"), 
    ) %>% 
  transmute(Group = group, n, Proficiency, Use, Proficiency_prose, Use_prose) %>% 
  mutate(Group = str_to_upper(Group)) %>% 
  write_csv(here("tables", "tab_prof_use_desc.csv"))

tab_prof_use_desc %>% 
  select(-c("Proficiency_prose", "Use_prose")) %>% 
  knitr::kable(format = "pandoc", 
    align = c("l", "r", "r", "r"), 
    caption = "Average use and proficiency for L2 and heritage learners.")

tab_use_prof_desc_mod <- use_prof_desc %>% 
  mutate(contrast = str_replace_all(contrast, "hl", "hs")) %>% 
  select(metric:out) %>% 
  separate(out, into = c("Estimate", "ROPE"), sep = "\n") %>% 
  separate(ROPE, into = c("ROPE", "PD"), sep = ",") %>% 
  mutate(contrast = str_to_upper(contrast), 
    Estimate = str_remove(Estimate, "β = "), 
    Estimate = str_replace_all(Estimate, "-", "\U2212"), 
    PD = str_remove(PD, "PD = "), 
    ROPE = str_remove(ROPE, "ROPE = "), 
    PD = str_remove(PD, " ")) %>% 
  select(Metric = metric, Contrast = contrast, Estimate, ROPE, PD) %>% 
  write_csv(here("tables", "tab_prof_use_mod.csv"))

tab_use_prof_desc_mod %>% 
  knitr::kable(format = "pandoc", 
    align = c("l", "l", "r", "r", "r"), 
    caption = "Marginal mean difference beween L2 and heritage leareners for use and proficiency.")
```

```{r}
#| label: save-final-use-prof-dataframe
prof_use %>% 
  mutate(remove = if_else(id %in% remove, 1, 0)) %>% 
  write_csv(here("data", "tidy", "prof_use.csv"))
```

