---
title: "{{< var title >}}: {{< var version >}}"
subtitle: "Eye-tracking data clean-up"
author: "{{< var author.name >}}"
date: "`r Sys.Date()`"
format: html
---

```{r}
#| label: load-packages
source(here::here("scripts", "00_libraries.R"))
source(here::here("scripts", "01_helpers.R"))
```

```{r}
#| label: load-data
# Load data and fix accent mark chars
stress_50_temp <- read_tsv(here("data", "raw", "stress_extract_50.txt")) %>% 
  mutate(target = str_replace(target, "ó", "o"))
```

Important variable info: 

- AVERAGE_IA_1_SAMPLE_COUNT: looks at target
- AVERAGE_IA_2_SAMPLE_COUNT: looks at distractor
- AVERAGE_IA_0_SAMPLE_COUNT: looks elsewhere

How much data do we lose by selecting only accurate trials: `r sum((stress_50_temp$ACCURACY == 0) / length(stress_50_temp$ACCURACY))`.


Now we take a look at item lengths and check some landmarks. 

```{r}
#| label: calculate-item-lengths
itemlength <- stress_50_temp %>%
  select(., target, t01, t04, t05, t06) %>%
  distinct() %>% 
  mutate(
    syll1_dur = t04 - t01, 
    syll2_dur = t06 - t04, 
    C3_dur    = t05 - t04, 
    item_dur  = t06 - t01, 
    derived_dur = syll1_dur + syll2_dur, 
    check = item_dur == derived_dur
  )

knitr::kable(itemlength %>% head, format = "pandoc")
```


Ok, let's tidy and center the time course.

```{r}
#| label: tidy-center-timecourse
stress_50_hold <- stress_50_temp %>%
  
  # create id variable
  rename(id = RECORDING_SESSION_LABEL) %>% 
  
  # create group variable
  separate(id, into = c("group", "trash"), sep = 3, remove = F) %>%
  
  #select and rename variables of interest
  dplyr::select(id, trial = TRIAL_INDEX, bin = BIN_INDEX,
    target_count = AVERAGE_IA_1_SAMPLE_COUNT, 
    target_prop = `AVERAGE_IA_1_SAMPLE_COUNT_%`,
    distractor_count = AVERAGE_IA_2_SAMPLE_COUNT, 
    distractor_prop = `AVERAGE_IA_2_SAMPLE_COUNT_%`, 
    elsewhere_count = AVERAGE_IA_0_SAMPLE_COUNT, 
    elsewhere_prop = AVERAGE_IA_0_SAMPLE_COUNT, 
    accuracy = ACCURACY, rt = RT, block, stress = cond, 
    offset_prev_word = t01, 
    onset_v1 = t02, 
    onset_c2 = t03, 
    onset_c3 = t04, 
    onset_v2 = t05, 
    offset_target = t06, 
    end_sentence = t07, target, version, group) %>%

  # remove incorrect
  filter(., accuracy == 1) %>%

  # Create eLog variable and respective wts
  mutate(eLog = log((target_count + 0.5) / (50 - target_count + 0.5)),
         wts = 1 / (target_count + 0.5) + 1 / (50 - target_count + 0.5)) %>%
    
  # Select necessary columns
  dplyr::select(id, group, target, stress, bin,
    target_count, target_prop, eLog, wts, onset_c3, 
    elsewhere_count, elsewhere_prop, distractor_count, distractor_prop) %>% 

  # Get suffix onset label and center at 0 for each
  # participant for each item
  # change 'onset_c3' in previous `select` call to center around diff. landmark
  mutate(lm_bin = (onset_c3 / 50) %>% ceiling(.),
         t_onset = if_else(bin == lm_bin, TRUE, FALSE)) %>% 
  group_by(id, target) %>%
  mutate(time_zero = center_timecourse(bin, t_onset, event = "TRUE")) %>% 
  ungroup()
```



Checks and fixes

```{r}
#| label: checks-fixes
#| out.width: "100%"
#| fig.height: 3.5

# Remove unwanted groups
stress_50 <- stress_50_hold %>% 
  filter(!(group %in% c("ams", "ims"))) %>% 
  mutate(id = fct_recode(id, ihs02 = "ih02"), 
    group = fct_recode(group, ihs = "ih0"), 
    group = case_when(
      group %in% c("ihs", "ahs") ~ "HL", 
      group %in% c("ies", "aes") ~ "L2", 
      TRUE ~ "Monolingual"), 
    stress = if_else(stress == 2, "Oxytone", "Paroxytone")
    ) 

# N x group
stress_50 %>% 
  group_by(group) %>% 
  summarize(n = n_distinct(id))

# Plot check
stress_50 %>% 
  filter(time_zero > -6 & time_zero < 11) %>% 
  mutate(stress = fct_relevel(stress, "Paroxytone")) %>% 
  ggplot(., aes(x = time_zero, y = target_prop)) + 
    facet_grid(. ~ stress) + 
    geom_hline(yintercept = 0.5, lty = 3) + 
    geom_vline(xintercept = 0 + 4, lty = 3) + 
    stat_summary(aes(fill = group), fun.data = mean_cl_boot, geom = "ribbon", 
      alpha = 0.5) + 
    stat_summary(aes(color = group), fun = mean, geom = "line") + 
    scale_color_viridis_d(name = NULL, end = 0.8) + 
    scale_fill_viridis_d(name = NULL, end = 0.9) + 
    coord_cartesian(xlim = c(NA, 10), ylim = c(-0.05, 1.05), expand = F) + 
    scale_x_continuous(breaks = seq(-4, 12, 4), labels = seq(-200, 600, 200)) + 
    labs(y = "P(target fixation)", x = "Time course") + 
    theme(legend.position = c(0.1, 0.86)) + 
    NULL

```

Looks ok. 
Now we add proficiency and use data for the bilingual groups, then we save the complete data frame and split it by ID so it will fit on github.

```{r}
#| label: combine-save-data

# Load prof/use data
prof_use <- read_csv(here("data", "tidy", "prof_use.csv")) %>% 
  select(id, prof, use, remove) %>% 
  filter(remove == 0) %>% 
  mutate(prof_z = (prof - mean(prof)) / sd(prof), 
    use_z = (use - mean(use)) / sd(use))

# Get vector of IDs to keep
keep_ids <- c(
  prof_use$id %>% unique, 
  filter(stress_50, group == "Monolingual") %>% pull(id) %>% 
    unique %>% as.character
)

# Check n per group to see if filter is correct
# Should be 42 HL and 50 L2
stress_50 %>% 
  filter(id %in% keep_ids) %>% 
  group_by(group) %>% 
  summarize(n = n_distinct(id))

# Join data frames and save
if(F) {
left_join(stress_50 %>% filter(id %in% keep_ids), prof_use, by = "id") %>%
  select(id:group, prof:use, prof_z:use_z, target:time_zero) %>% 
  write_csv(here("data", "tidy", "stress_50ms.csv")) %>% 
  group_by(id) %>% 
  group_walk(~ write_csv(.x, 
    here("data", "tidy", "by_id_50ms", 
      paste0(.y$id, "_stress_50ms.csv"))))
}
```
